---
params:
  pid: "Patient"
  sid: "Sample"
  mid: ""
  date: "`r Sys.time()`"
  marker.channel: 1
  alb.channel: 2
  dx.alb: ""
  fu.alb: ""
  mnc.alb: ""
  h2o.alb: ""
  dx.marker: ""
  mnc.marker: ""
  h2o.marker: ""
  fu.marker: ""
  dx.baseline: NA
  mnc.alb.dilutionX: 200
  mnc.alb.concentration: ""
  show_outliers: TRUE
title: "ddPCR laboratory results"
fontsize: 10pt
geometry: margin=1.5cm
fig_width: 8 
fig_height: 6
output:
  pdf_document:
    keep_tex: false
    toc: false
    highlight: zenburn
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lastpage}
---


<!--- Define Headers and Footers --->
\fancypagestyle{plain}{%
  \renewcommand{\headrulewidth}{0pt}%
  \fancyhf{}%
  \fancyfoot[C]{\footnotesize Page \thepage\, of\, \pageref*{LastPage} `r params$sid` `r params$mid` Generated at `r params$date`}
  \setlength\footskip{0pt}
}
\pagestyle{plain}
\center
# Sample `r params$sid`
\center
# Marker `r params$mid`


```{r echo=FALSE}
library(knitr)
plot1d <- function(multiwell.results, channel, ...) {
	x <- c()
	y <- c()
	mask <- c()
	seg <- c(0)
	lower <- c()
	higher <- c()
	mids <- c()
	col <- c()
	current.x <- 0
	for (results in multiwell.results) {
		x <- c(x, current.x + (1:nrow(results$intensities)))
		
		y <- c(y, results$intensities[, channel])
		mask <- c(mask, results$mask)
		lower <- c(lower, current.x)
		mids <- c(mids, current.x + 0.5 * nrow(results$intensities))
		current.x <- current.x + nrow(results$intensities)
		higher <- c(higher, current.x)
		col <- c(col, 1 + (results$intensities[, channel] > results$threshold))
		seg <- c(seg, current.x + 0.5)
	}
	plot(x, y, cex=1, xlab="", ylab = "intensities", axes=FALSE, pch=c(4, 16)[mask + 1], col=col, ...)
	box()
	axis(2)
	axis(1, at = mids, labels = FALSE)
	
	text(x = mids,
     y = par("usr")[3] - (max(y) - min(y))/30,
     labels = sapply(multiwell.results, "[[", "name"),
     xpd = NA,
     srt = 25,
     cex = 0.8, adj=1)
	 
	for (i in 1:length(multiwell.results)) {
		lines(c(lower[i], higher[i]), c(multiwell.results[[i]]$threshold, multiwell.results[[i]]$threshold), lty=1, col=2, lwd=2)
	}
	for (i in 1:length(seg)) {
		lines(c(seg[i], seg[i]), c(1e-6, 1e6), col="blue", lty=1)
	}
}
plot2d <- function(singlewell.results, channel, ...) {
		plot(singlewell.results$intensities[, marker.channel], singlewell.results$intensities[, alb.channel], col = (singlewell.results$intensities[, channel] > singlewell.results$threshold) + 1,
			pch=c(4, 16)[singlewell.results$mask + 1],
			main=paste(singlewell.results$name, names(singlewell.results$name)), xlab = "Channel 1", ylab="Channel 2", ...)
		if (channel == 1) {
			lines(c(singlewell.results$threshold, singlewell.results$threshold), c(-1e8, 1e8), col=2, lwd=1.5)
		} else if (channel == 2) {
			lines(c(-1e8, 1e8), c(singlewell.results$threshold, singlewell.results$threshold), col=2, lwd=1.5)
		}
}
get_info <- function(multiwell.results) {
	tab <- t(sapply(multiwell.results, function(results) {
		events <- nrow(results$intensities)
		outliers <- results$n.outliers
		used.events <- events - outliers
		threshold <- results$threshold
		positive <- results$n.positive.droplets
		concentration <- results$concentration
		c(events, outliers, used.events, threshold, positive, concentration)
	}))
	rownames(tab) <- sapply(multiwell.results, "[[", "name")
	colnames(tab) <- c("events", "outliers", "used.events", "threshold", "positive", "concentration")
	tab
}

get_merged_info <- function(multiwell.results, sample.names=NA) {
	if (is.na(sample.names)) {
		sample.names = sapply(multiwell.results, "[[", "name")
	
	}
	tab <- get_info(multiwell.results)
	unique.names <- unique(sample.names)
	
	total.events <- tapply(tab[, "events"], sample.names, sum)[unique.names]
	total.outliers <- tapply(tab[, "outliers"], sample.names, sum)[unique.names]
	total.used.events <- tapply(tab[, "used.events"], sample.names, sum)[unique.names]
	total.positive <- tapply(tab[, "positive"], sample.names, sum)[unique.names]
	concentration <- -log(1-(total.positive/total.used.events))/0.00085
	concentration <- concentration[unique.names]
	replicates <- tapply(tab[, "events"], sample.names, length)[unique.names]
	info <- cbind(replicates=replicates, events=total.events, outliers=total.outliers, used.events=total.used.events,positive=total.positive,concentration=concentration )
	rownames(info) <- unique.names
	info
}
```



# Albumin Results

```{r fig1, fig.height=7, fig.width=12, fig.align="center", echo=FALSE}
plot1d(c(params$dx.alb, params$mnc.alb, params$h2o.alb, params$fu.alb), 2, main="Intensities and annotation of albumin wells")
alb.info <- get_info(c(params$dx.alb, params$mnc.alb, params$h2o.alb, params$fu.alb))
alb.merged.info <- get_merged_info(c(params$dx.alb, params$mnc.alb, params$h2o.alb, params$fu.alb))
kable(alb.info, caption = "Per well summary of Albumin results")
kable(alb.merged.info, caption = "Per sample summary of Albumin results")
```

# Marker Results

```{r fig2, fig.height=7, fig.width=12, fig.align="center", echo=FALSE}
plot1d(c(params$dx.marker, params$mnc.marker, params$h2o.marker, params$fu.marker),1, main="Intensities and annotation of marker wells")
marker.info <- get_info(c(params$dx.marker, params$mnc.marker, params$h2o.marker, params$fu.marker))
marker.merged.info <- get_merged_info(c(params$dx.marker, params$mnc.marker, params$h2o.marker, params$fu.marker))
kable(marker.info, caption = "Per well summary of marker results")
kable(marker.merged.info, caption = "Per sample summary of marker results")
```

# MRD calculation

```{r mrd_calculation, echo=FALSE}
# cell contration at fullow up
fu.alb.info <- get_merged_info(c(params$fu.alb))
fu.alb.dilution.factor <- as.numeric(gsub("x$", "", strsplit(rownames(fu.alb.info), "_", fixed=TRUE)[[1]][4]))
fu.cell.concentration <- fu.alb.info[1, "concentration"] * fu.alb.dilution.factor / 2
if (nrow(fu.alb.info) > 1) {
	print("warning, there are more than 1 albumin sample at follow up.")
}
# marker concentration at follow up
fu.marker.info <- get_merged_info(c(params$fu.marker))
# absolute mrd
absolute.mrd <- fu.marker.info[1, "concentration"] / fu.cell.concentration

# diagnostic baseline
dx.baseline <- params$dx.baseline
if (is.na(dx.baseline)) {
	dx.alb.info <- get_merged_info(c(params$dx.alb))
	dx.cell.concentration <- dx.alb.info[1, "concentration"] / 2
	dx.marker.info <- get_merged_info(c(params$dx.marker))
	dx.baseline <- dx.marker.info[1, "concentration"] / dx.cell.concentration
	writeLines(paste("Using in-plate diagnostic baseline."))
} else {
	writeLines(paste("Using preset diagnostic baseline", dx.baseline))
}

# relatiive mrd
relative.mrd <- absolute.mrd/dx.baseline

# mrd in mnc
mnc.marker.info <- get_merged_info(c(params$mnc.marker))

if (mnc.marker.info[1, "positive"] == 0) {
	mnc.marker.info[1, "positive"] <- 1
	mnc.marker.info[1, "concentration"] <- -log(1-(mnc.marker.info[1, "positive"]/mnc.marker.info[1, "used.events"]))/0.00085
	writeLines("No positive event of marker observed in MNC wells. 1 positive event assumed.")
}

mnc.alb.info <- get_merged_info(c(params$mnc.alb))
mnc.alb.dilution.factor <- as.numeric(gsub("x$", "", strsplit(rownames(mnc.alb.info), "_", fixed=TRUE)[[1]][3]))

#background.level <- mnc.marker.info[1, "concentration"]*2/(mnc.alb.info[1, "concentration"] * mnc.alb.dilution.factor)
background.level <- mnc.marker.info[1, "concentration"]/(params$mnc.alb.concentration * params$mnc.alb.dilutionX / 2)

# total cell no.
total.cell.no <- ceiling(fu.cell.concentration * 20 * fu.marker.info[1, "replicates"])

# thresholds
ptv <- qbinom(0.9999, total.cell.no, background.level)/(20 * fu.marker.info[1, "replicates"])
ntv <- qbinom(0.95, total.cell.no, background.level)/(20 * fu.marker.info[1, "replicates"])
# mrd thresholds
mrd.ptv <- (ptv/fu.cell.concentration)/dx.baseline
mrd.ntv <- (ntv/fu.cell.concentration)/dx.baseline

#
mrd.call <- "NEG"
if (relative.mrd > mrd.ptv) {
	mrd.call <- "POS"
} else if (relative.mrd > mrd.ntv) {
	mrd.call <- "LOQ"
}


fu.marker.formatted <- sprintf("%#.3f", fu.marker.info[1, "concentration"])

fu.cell.concentration.formatted <- sprintf("%#.3f", fu.cell.concentration)

absolute.mrd.formatted <- formatC(absolute.mrd, format = "e", digits = 2)


dx.baseline.formatted <- sprintf("%#.3f", dx.baseline)

background.level.formatted <- formatC(background.level, format = "e", digits = 2)

relative.mrd.formatted <- formatC(relative.mrd, format = "e", digits = 2)
mrd.ntv.formatted <- formatC(mrd.ntv, format = "e", digits = 2)
mrd.ptv.formatted <- formatC(mrd.ptv, format = "e", digits = 2)


mrd.output.values <- c(fu.marker.formatted, fu.cell.concentration.formatted, absolute.mrd.formatted,  dx.baseline.formatted, background.level.formatted, relative.mrd.formatted, total.cell.no,  mrd.ntv.formatted, mrd.ptv.formatted, mrd.call)

names(mrd.output.values) <- c("FU Marker copy/uL", "FU Alb copy/uL", "Absolute MRD", "Dx Baseline", "MNC Baseline", "MRD", "Total Cell Number", "NTV", "PTV", "Call")

mrd.output.values <- as.matrix(mrd.output.values)

colnames(mrd.output.values) <- "Values"

kable(mrd.output.values, caption = "Summary of MRD calculation.")

```

# 2D Figures

## Albumin, diagnostic sample

```{r 2dFigures_dx_alb, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$dx.alb)) {
	plot2d(params$dx.alb[[i]], alb.channel)
	#info <- get_info(params$dx.alb[i])
	#kable(info, caption = paste("Summary of", params$dx.alb[[i]]$name))
}
```


## Albumin, MNC

```{r 2dFigures_mnc_alb, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$mnc.alb)) {
	plot2d(params$mnc.alb[[i]], alb.channel)
	#info <- get_info(params$mnc.alb[i])
	#kable(info, caption = paste("Summary of", params$mnc.alb[[i]]$name))
}
```


## Albumin, NTC/H2O

```{r 2dFigures_h2o_alb, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
results <- c(params$mnc.alb, params$h2o.alb)
marker.channel.max <- max(sapply(results, function(x) {max(x$intensities[, marker.channel])}))
marker.channel.min <- min(sapply(results, function(x) {min(x$intensities[, marker.channel])}))
alb.channel.max <- max(sapply(results, function(x) {max(x$intensities[, alb.channel])}))
alb.channel.min <- min(sapply(results, function(x) {min(x$intensities[, alb.channel])}))


for (i in 1:length(params$h2o.alb)) {
	plot2d(params$h2o.alb[[i]], alb.channel, xlim=c(marker.channel.min, marker.channel.max), ylim=c(alb.channel.min, alb.channel.max))
	#info <- get_info(params$h2o.alb[i])
	#kable(info, caption = paste("Summary of", params$h2o.alb[[i]]$name))
}
```



## Albumin, follow up

```{r 2dFigures_fu_alb, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$fu.alb)) {
	plot2d(params$fu.alb[[i]], alb.channel)
	#info <- get_info(params$fu.alb[i])
	#kable(info, caption = paste("Summary of", params$fu.alb[[i]]$name))
}
```




## Marker, diagnostic sample

```{r 2dFigures_dx_marker, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
marker.results <- c(params$dx.marker, params$mnc.marker, params$h2o.marker, params$fu.marker)
marker.channel.max <- max(sapply(marker.results, function(x) {max(x$intensities[, marker.channel])}))
marker.channel.min <- min(sapply(marker.results, function(x) {min(x$intensities[, marker.channel])}))
alb.channel.max <- max(sapply(marker.results, function(x) {max(x$intensities[, alb.channel])}))
alb.channel.min <- min(sapply(marker.results, function(x) {min(x$intensities[, alb.channel])}))


for (i in 1:length(params$dx.marker)) {
	plot2d(params$dx.marker[[i]], marker.channel, xlim=c(marker.channel.min, marker.channel.max), ylim=c(alb.channel.min, alb.channel.max))
	#info <- get_info(params$dx.marker[i])
	#kable(info, caption = paste("Summary of", params$dx.marker[[i]]$name))
}
```


## Marker, MNC

```{r 2dFigures_mnc_marker, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$mnc.marker)) {
	plot2d(params$mnc.marker[[i]], marker.channel, xlim=c(marker.channel.min, marker.channel.max), ylim=c(alb.channel.min, alb.channel.max))
	#info <- get_info(params$mnc.marker[i])
	#kable(info, caption = paste("Summary of", params$mnc.marker[[i]]$name))
}
```


## Marker, NTC/H2O

```{r 2dFigures_h2o_marker, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$h2o.marker)) {
	plot2d(params$h2o.marker[[i]], marker.channel, xlim=c(marker.channel.min, marker.channel.max), ylim=c(alb.channel.min, alb.channel.max))
	#info <- get_info(params$h2o.marker[i])
	#kable(info, caption = paste("Summary of", params$h2o.marker[[i]]$name))
}
```



## Marker, follow up

```{r 2dFigures_fu_marker, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$fu.marker)) {
	plot2d(params$fu.marker[[i]], marker.channel, xlim=c(marker.channel.min, marker.channel.max), ylim=c(alb.channel.min, alb.channel.max))
	#info <- get_info(params$fu.marker[i])
	#kable(info, caption = paste("Summary of", params$fu.marker[[i]]$name))
}
```

