---
params:
  pid: "Patient"
  sid: "Sample"
  mid: ""
  date: "`r Sys.time()`"
  marker.channel: 1
  alb.channel: 2
  dx.alb: ""
  fu.alb: ""
  mnc.alb: ""
  h2o.alb: ""
  dx.marker: ""
  mnc.marker: ""
  h2o.marker: ""
  fu.marker: ""
  dx.baseline: NA
  mnc.alb.dilutionX: 200
  mnc.alb.concentration: ""
  show_outliers: TRUE
title: "ddPCR laboratory results - 2D Figures"
fontsize: 10pt
geometry: margin=1.5cm
fig_width: 8 
fig_height: 6
output:
  pdf_document:
    keep_tex: false
    toc: false
    highlight: zenburn
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lastpage}
---


<!--- Define Headers and Footers --->
\fancypagestyle{plain}{%
  \renewcommand{\headrulewidth}{0pt}%
  \fancyhf{}%
  \fancyfoot[C]{\footnotesize Page \thepage\, of\, \pageref*{LastPage} `r params$sid` `r params$mid` Generated at `r params$date` by ddPCR MRD Pipeline Version 1.0}
  \setlength\footskip{0pt}
}
\pagestyle{plain}
\center
# Sample `r params$sid`, Marker `r params$mid`


```{r echo=FALSE}
library(knitr)
plot1d <- function(multiwell.results, channel, ...) {
	x <- c()
	y <- c()
	mask <- c()
	seg <- c(0)
	lower <- c()
	higher <- c()
	mids <- c()
	col <- c()
	current.x <- 0
	for (results in multiwell.results) {
		x <- c(x, current.x + (1:nrow(results$intensities)))
		
		y <- c(y, results$intensities[, channel])
		mask <- c(mask, results$mask)
		lower <- c(lower, current.x)
		mids <- c(mids, current.x + 0.5 * nrow(results$intensities))
		current.x <- current.x + nrow(results$intensities)
		higher <- c(higher, current.x)
		col <- c(col, 1 + (results$intensities[, channel] > results$threshold))
		seg <- c(seg, current.x + 0.5)
	}
	plot(x, y, cex=1, xlab="", ylab = "intensities", axes=FALSE, pch=c(4, 16)[mask + 1], col=col, ...)
	box()
	axis(2)
	axis(1, at = mids, labels = FALSE)
	
	text(x = mids,
     y = par("usr")[3] - (max(y) - min(y))/30,
     labels = sapply(multiwell.results, "[[", "name"),
     xpd = NA,
     srt = 25,
     cex = 0.8, adj=1)
	 
	for (i in 1:length(multiwell.results)) {
		lines(c(lower[i], higher[i]), c(multiwell.results[[i]]$threshold, multiwell.results[[i]]$threshold), lty=1, col=2, lwd=2)
	}
	for (i in 1:length(seg)) {
		lines(c(seg[i], seg[i]), c(1e-6, 1e6), col="blue", lty=1)
	}
}
plot2d <- function(singlewell.results, channel, ...) {
		plot(singlewell.results$intensities[, marker.channel], singlewell.results$intensities[, alb.channel], col = (singlewell.results$intensities[, channel] > singlewell.results$threshold) + 1,
			pch=c(4, 16)[singlewell.results$mask + 1],
			main=paste(singlewell.results$name, names(singlewell.results$name)), xlab = "Channel 1", ylab="Channel 2", ...)
		if (channel == 1) {
			lines(c(singlewell.results$threshold, singlewell.results$threshold), c(-1e8, 1e8), col=2, lwd=1.5)
		} else if (channel == 2) {
			lines(c(-1e8, 1e8), c(singlewell.results$threshold, singlewell.results$threshold), col=2, lwd=1.5)
		}
}
get_info <- function(multiwell.results) {
	tab <- t(sapply(multiwell.results, function(results) {
		events <- nrow(results$intensities)
		outliers <- results$n.outliers
		used.events <- events - outliers
		threshold <- results$threshold
		positive <- results$n.positive.droplets
		concentration <- results$concentration
		c(events, outliers, used.events, threshold, positive, concentration)
	}))
	rownames(tab) <- sapply(multiwell.results, "[[", "name")
	colnames(tab) <- c("events", "outliers", "used.events", "threshold", "positive", "concentration")
	tab
}

get_merged_info <- function(multiwell.results, sample.names=NA) {
	if (is.na(sample.names)) {
		sample.names = sapply(multiwell.results, "[[", "name")
	
	}
	tab <- get_info(multiwell.results)
	unique.names <- unique(sample.names)
	
	total.events <- tapply(tab[, "events"], sample.names, sum)[unique.names]
	total.outliers <- tapply(tab[, "outliers"], sample.names, sum)[unique.names]
	total.used.events <- tapply(tab[, "used.events"], sample.names, sum)[unique.names]
	total.positive <- tapply(tab[, "positive"], sample.names, sum)[unique.names]
	concentration <- -log(1-(total.positive/total.used.events))/0.00085
	concentration <- concentration[unique.names]
	replicates <- tapply(tab[, "events"], sample.names, length)[unique.names]
	info <- cbind(replicates=replicates, events=total.events, outliers=total.outliers, used.events=total.used.events,positive=total.positive,concentration=concentration )
	rownames(info) <- unique.names
	info
}
```

## Albumin, diagnostic sample

```{r 2dFigures_dx_alb, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$dx.alb)) {
	plot2d(params$dx.alb[[i]], alb.channel)
	#info <- get_info(params$dx.alb[i])
	#kable(info, caption = paste("Summary of", params$dx.alb[[i]]$name))
}
```


## Albumin, MNC

```{r 2dFigures_mnc_alb, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$mnc.alb)) {
	plot2d(params$mnc.alb[[i]], alb.channel)
	#info <- get_info(params$mnc.alb[i])
	#kable(info, caption = paste("Summary of", params$mnc.alb[[i]]$name))
}
```


## Albumin, NTC/H2O

```{r 2dFigures_h2o_alb, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
results <- c(params$mnc.alb, params$h2o.alb)
marker.channel.max <- max(sapply(results, function(x) {max(x$intensities[, marker.channel])}))
marker.channel.min <- min(sapply(results, function(x) {min(x$intensities[, marker.channel])}))
alb.channel.max <- max(sapply(results, function(x) {max(x$intensities[, alb.channel])}))
alb.channel.min <- min(sapply(results, function(x) {min(x$intensities[, alb.channel])}))


for (i in 1:length(params$h2o.alb)) {
	plot2d(params$h2o.alb[[i]], alb.channel, xlim=c(marker.channel.min, marker.channel.max), ylim=c(alb.channel.min, alb.channel.max))
	#info <- get_info(params$h2o.alb[i])
	#kable(info, caption = paste("Summary of", params$h2o.alb[[i]]$name))
}
```



## Albumin, follow up

```{r 2dFigures_fu_alb, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$fu.alb)) {
	plot2d(params$fu.alb[[i]], alb.channel)
	#info <- get_info(params$fu.alb[i])
	#kable(info, caption = paste("Summary of", params$fu.alb[[i]]$name))
}
```




## Marker, diagnostic sample

```{r 2dFigures_dx_marker, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
marker.results <- c(params$dx.marker, params$mnc.marker, params$h2o.marker, params$fu.marker)
marker.channel.max <- max(sapply(marker.results, function(x) {max(x$intensities[, marker.channel])}))
marker.channel.min <- min(sapply(marker.results, function(x) {min(x$intensities[, marker.channel])}))
alb.channel.max <- max(sapply(marker.results, function(x) {max(x$intensities[, alb.channel])}))
alb.channel.min <- min(sapply(marker.results, function(x) {min(x$intensities[, alb.channel])}))


for (i in 1:length(params$dx.marker)) {
	plot2d(params$dx.marker[[i]], marker.channel, xlim=c(marker.channel.min, marker.channel.max), ylim=c(alb.channel.min, alb.channel.max))
	#info <- get_info(params$dx.marker[i])
	#kable(info, caption = paste("Summary of", params$dx.marker[[i]]$name))
}
```


## Marker, MNC

```{r 2dFigures_mnc_marker, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$mnc.marker)) {
	plot2d(params$mnc.marker[[i]], marker.channel, xlim=c(marker.channel.min, marker.channel.max), ylim=c(alb.channel.min, alb.channel.max))
	#info <- get_info(params$mnc.marker[i])
	#kable(info, caption = paste("Summary of", params$mnc.marker[[i]]$name))
}
```


## Marker, NTC/H2O

```{r 2dFigures_h2o_marker, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$h2o.marker)) {
	plot2d(params$h2o.marker[[i]], marker.channel, xlim=c(marker.channel.min, marker.channel.max), ylim=c(alb.channel.min, alb.channel.max))
	#info <- get_info(params$h2o.marker[i])
	#kable(info, caption = paste("Summary of", params$h2o.marker[[i]]$name))
}
```



## Marker, follow up

```{r 2dFigures_fu_marker, fig.height=12, fig.width=12, fig.align="center", echo=FALSE}
for (i in 1:length(params$fu.marker)) {
	plot2d(params$fu.marker[[i]], marker.channel, xlim=c(marker.channel.min, marker.channel.max), ylim=c(alb.channel.min, alb.channel.max))
	#info <- get_info(params$fu.marker[i])
	#kable(info, caption = paste("Summary of", params$fu.marker[[i]]$name))
}
```

